### 2021-05-21 docs/tapf.tex

\documentclass[a4paper,10pt,english]{article}
\usepackage[T1]{fontenc}
\usepackage{marginnote}
\include{prelude}
\usepackage{geometry}
\geometry{verbose,tmargin=3cm,bmargin=3cm,lmargin=3cm,rmargin=3cm}
\usepackage[backend=bibtex]{biblatex}
\addbibresource{sources.bib}
\usepackage{babel}

\title{Reviewed: Train Unit Shunting and Servicing: a Real-Life Application of Multi-Agent Path Finding}
\author{Thom van der Woude}
\date{\today}

\begin{document}
	\maketitle
	\paragraph{First run-through}
	\[\mathcal{O}(k_{t_{max}}!)\]
	\[t_{max}\]
	\[k_{t_{max}}\]
	\printbibliography

\end{document}